\subsection{Content References from the Pre-Existing Term Project}
The following sections originated in the pre-existing semester thesis "Dafny Server Redesign" \cite{sa} and were re-included into this document for the sense of a comprehensive documentation.
Minor changes, such as typos or the adjustment of the preceding project, are not mentioned separately.\\

Dafny and the language server protocol are fundamental technologies of this project.
Our supervisors were following the development of the preceding semester thesis and are familiar with these technologies.
However, the external expert, as well as the third reader were not involved.
To provide them with a comprehensive understanding of all necessary context, the analysis of those technologies done in the preceding thesis was copied into this document.

\intnote{und sonst noch: alle aussagen brauchen quellen todo}


\begin{itemize}
    \item Chapter \ref{section:abstract} \nameref{section:abstract}
        \begin{itemize}
            \item Paragraph 1 about Dafny
            \item Paragraph 2 about the language server protocol communication
        \end{itemize}

    \item Chapter \ref{section:management_summary} \nameref{section:management_summary}
        \begin{itemize}
            \item Section \ref{section:management_summary:dafny} about Dafny
            \item Section \ref{section:management_summary:lsp} about the language server protocol communication
        \end{itemize}

    \item Chapter \ref{section:introduction} \nameref{section:introduction}
        \begin{itemize}
            \item Chapter \ref{section:introduction:dafny} except for the paragraph about lemmas
            \item Chapter \ref{section:introduction:initialsolution}, partially
        \end{itemize}

    \item Chapter \ref{section:analysis} \nameref{section:analysis}
        \begin{itemize}
            \item Chapter \ref{section:analysis:lsp}
            \item Chapter \ref{section:analysis:omnisharp}, reworked, better example
            \item Chapter \ref{section:analysis:features}, adoption of the starting positions from the results. How the improvements are to be implemented is new.
        \end{itemize}

    \item Chapter \ref{section:results} \nameref{section:results}
        \begin{itemize}
            \item Chapter \ref{section:result:syntaxhighgliht}
        \end{itemize}

    \item Chapter \ref{section:project_management} \nameref{section:project_management}
        \begin{itemize}
            \item The whole chapter is structurally based on the pre-existing term project.
            Partially, sentence fragments from general introductions were taken over.
            However, the data as well as the conclusions from it are new.
        \end{itemize}


\end{itemize}
